### 2016-01-13 model/overflowmcmc.tex
Priors and algorithm sketch finished

\subsection{Priors on $\alpha_d$ and $\beta_d$}

\subsection{Priors on $\bar{D}_k$'s}

Finally, the upper limits are modeled as a categorical distribution over a set of prespecified possible limits, $\Delta_1,\dots,\Delta_{n_{lim}}$ with probability $p_{ik} = P(\bar{D}_k = \Delta_i)$, $\sum_ip_{ik} = 1$ and $p_{ik} \ge 0$. In other words
\begin{align*}
\bar{D}_k\stackrel{ind}{\sim} Cat(\Delta_{1:n_{lim}}, p_{1:n_{lim},k}).
\end{align*}

@@

\subsection{Block arrival times --- $\tau_k$}
The full conditional distribution of the block arrival times is more complicated:
\begin{align*}
p(\bm{\tau}_{1:n_{b}}|...) \propto \exp\left[-\frac{1}{2\sigma^2}\sum_{k=1}^{n_{b}}\sum_{i=1}^{\tilde{n}_k+1}\left(\tilde{t}_{ik} - \tau_k\right)^2 - \frac{1}{10}\tau_{n_{b}}\right]\frac{\phi^{\Psi(\tau_k) - \Psi(\tau_{k-1})}}{\Gamma\left[\Psi(\tau_k) - \Psi(\tau_{k-1})\right]}\lambda_k^{\Psi(\tau_k)-\Psi(\tau_{k-1}) - 1}
\end{align*}
where $\tau_0\equiv0$ and $\Gamma(x) = \int_0^\infty t^{x-1}e^{-t}dt$ is the gamma function. Even if $\Psi(\tau)$ is some nice known form, this density is nonstandard because the $\tau_k$'s enter the Gamma function. As a result, we are limited to several options for this step:
\begin{enumerate}
\item Draw each $\tau_k$ condtional on all the others using random walk Metropolis steps for $k=1,\dots,n_{b}$.
\item Approximate the density of $\tau_k$ sufficiently well and use independent Metropolis steps for $k=1,\dots,n_{b}$.
\item Approximate the density of $\tau_{1:n_{b}}$ sufficiently well and use an independent Metropolis step.
\end{enumerate}
Options 2 and 3 could both be replaced with a rejection sampling step under some circumstances, but a Metropolis step is likely cheaper computationally [CITE THE LIU PAPER].

Ideas:
\begin{enumerate}
\item ARMS -- ARS except with a Metropolis correction for non-log concave densities (look it up).
\item RW Metropolis and tune the proposal variances.
\end{enumerate}

@@

\subsection{Regression parameters --- $\bm{\beta}$}
The full conditional posterior of $\bm{\beta}$ is closely connected to the functional form of the covariates in time. No matter the functional form, however, the density is complex and a Metropolis step is likely warranted. Recall that since we defined $\psi(t) = e^{\bm{z}(t)'\bm{\beta}}$ we have $\psi_k(\bm{\beta}) = \int_{\tau_{k-1}}^{\tau^k}e^{\bm{z}(t)'\bm{\beta}}dt$ --- in this subsection we write $\psi_k(\bm{\beta})$ to explicitly acknowledge that $\psi_k$ is a function of $\bm{\beta}$. Then the full conditional of $\bm{\beta}$ is
\begin{align*}
p(\bm{\beta}|\dots) \propto p(\bm{\beta})\prod_{k=1}^{n_b}\frac{\phi^{\psi_k(\bm{\beta})}\lambda_k^{\psi_k(\bm{\beta}) - 1}}{\Gamma[\psi_k(\bm{\beta})]}
\end{align*}
where in the prior $\bm{\beta} \sim p(\bm{\beta})$ independent of the other parameters. This density is complex and nonstandard since $\psi_k(\bm{\beta})$ enters the gamma function.

@@

Otherwise, draw $\eta_k \sim Poi(\lambda_k)1(\eta_k \ge M_k - N_{k-1})$ and set $N_k = N_{k-1} + \eta_k$.
\end{enumerate}
\subsection{Block size caps --- $\bar{D}_k$}
Suppose for $k=1,\dots,n_b$ we have independent categorical priors on each of the $\bar{D}_k$'s. That is $\bar{D}_k \sim Cat(\bm{a}_{1:n_{cat}}, \bm{p}_{k})$ where $P(\bar{D}_k = a_i) = p_{ik}$ so that $p_{ik}\ge 0$ for $i=1,2,\dots,n_{cat}$ and $\sum_{i=1}^{n_{cat}}p_{ik} = 1$, both for $k=1,2,\dots,n_b$. In the posterior we learn about $\bar{D}_k$ through the observed $D_k$'s. The caps are then independent in their full conditional and for each block $k$, there are two possibilities:
\begin{enumerate}
\item When $N_k > M_k$, $D_k \le \bar{D}_k < D_k + d_{M_k+1}$.
\item When $N_k = M_k$, $D_k \le \bar{D}_k$.
\end{enumerate}
In the first case, this should completely determine the value of $\bar{D}_k$ since transaction sizes are much smaller than the differences between possible caps. For both cases we simple draw $\bar{D}_k$ from a categorical distribution where the possible categories are only those remaining which satisfy the relevant constraint fiven above, and the probability of obtaining each category is proportional to its probability in the prior. For example, suppose only $a_1$ and $a_2$ satisfy the relevant constraint for $\bar{D}_k$. Then we draw $\bar{D}_k$ with probabilities $P(\bar{D}_k = a_1) = p_{1k}/(p_{1k} + p_{2k})$ and $P(\bar{D}_k = a_2) = p_{2k}/(p_{1k} + p_{2k})$.
\subsection{Transaction data parameters --- $\alpha_d$ and $\beta_d$}
The full conditional distribution of $(\alpha_d,\beta_d)$ is 
\begin{align*}
p(\alpha_d,\beta_d|...) \propto p(\alpha_d)p(\beta_d) \prod_{i=1}^{n_d}\frac{\beta_d^{\alpha_d}}{\Gamma(\alpha_d)}d_i^{\alpha_d - 1}e^{-\beta_dd_i}.
\end{align*}
When $\beta \sim G(a_\beta, b_\beta)$ in the prior, its full conditional becomes
\begin{align*}
p(\beta_d|...) \propto \beta_d^{\alpha_dn_d + a_{\beta} - 1}e^{-\beta_d\left(\sum_{i=1}^{n_d}d_i + b_{\beta}\right)},
\end{align*}
which is another gamma distribution. The full conditional of $\alpha_d$ is complex no matter what its prior is since $\alpha_d$ enters the gamma function, and has the form
\begin{align*}
p(\alpha_d|...) \propto p(\alpha_d)\frac{\left(\prod_{i=1}^{n_d}d_i + \beta_d^{n_d}\right)^{\alpha_d}}{\Gamma(\alpha_d)^{n_d}}.
\end{align*}
This density can be drawn from using rejection sampling or an independence Metropolis step, e.g. ARS or ARMS.
\bibliographystyle{apalike} 
\bibliography{../paper/btc}

>>>>

\subsection{Priors on $\alpha_d$ and $\beta_d$}
The conditionally conjugate prior for $beta_d$ is a gamma while $\alpha_d$ has no convenient conditionally conjugate form. We will assume that $\beta_d \sim G(a_\beta, b_\beta)$ and $\alpha_d \sim G(a_\alpha, b_\alpha)$. In order to choose $a_\alpha$, $b_\alpha$, $a_\beta$, and $b_\beta$, consider that $E[d_i|\alpha_d,\beta_d] = \alpha_d/\beta_d$. But 
\begin{align*}
E\left[\frac{\alpha_d}{\beta_d}\right] = E[\alpha_d]E\left[\frac{1}{\beta_d}\right] = \frac{a_\alpha}{b_\alpha}\frac{b_\beta}{a_\beta - 1}.
\end{align*}
Similarly
\begin{align*}
V\left[\frac{\alpha_d}{\beta_d}\right] = V[\alpha_d] + V\left[\frac{1}{\beta_d}\right] = \frac{a_\alpha}{b_\alpha^2} + \frac{b_\beta^2}{(a_\beta - 1)^2(a_\beta - 2)}.
\end{align*}
We will choose these parameters so that $E[\alpha_d/\beta_d] \approx 0.01$ MB and $V[\alpha_d/\beta_d]$ is large so that the prior is weakly informative. We will set $a_\beta = 2.1$ to ensure the variance calculated above exists. Then $a_\alpha = 0.1$, $b_\alpha = 1$, and $b_\beta = 1$ yields
\begin{align*}
E\left[\frac{\alpha_d}{\beta_d}\right] = \frac{1}{21} \approx 0.09; && V\left[\frac{\alpha_d}{\beta_d}\right] \approx 10.
\end{align*}

\subsection{Priors on the $\bar{D}_k$'s}
We will put a discrete prior on each of the $\bar{D}_k$'s. Theoretically a miner can put any upper limit she wants on a mined bitcoin, but in practice most miners use one of several possible values. We will use the set of values $\mathcal{D} = \{0.25, 0.35, 0.5, 0.75, 0.9, 0.95, 1\}$ [DOUBLE CHECK THESE --- IN PARTICULAR, ARE THERE OTHER COMMON VALUES? WHAT ABOUT THE ELGIUS MINING POOL'S CAPS?]. For each $\bar{D}_k$ we will use a discrete uniform prior on $\mathcal{D}$ so that $P(\bar{D}_k = j) = 1/7$ for all $j \in \mathcal{D}$.

[OPTIONS FOR MORE INTERESTING THINGS HERE: DIFFERENT PRIOR FOR EACH $k$ SO THAT LATER BLOCKS HAVE A HIGHER PROBABILITY OF HAVING A HIGHER CAP. THIS USES THE DATA TWICE, SO A BETTER OPTION MAY JUST BE MODELING THE PROBABILITIES ON $\mathcal{D}$ AS EVOLVING OVER TIME BASED ON SOME HIGHER LEVEL PARAMETERS]

@@

\subsection{Block arrival times --- $\tau_k$}
The full conditional distribution of the block arrival times is more complicated:
\begin{align*}
p(\bm{\tau}_{1:n_{b}}|...) \propto \exp\left[-\frac{1}{2\sigma^2}\sum_{k=1}^{n_{b}}\sum_{i=1}^{\tilde{n}_k+1}\left(\tilde{t}_{ik} - \tau_k\right)^2 - \frac{1}{10}\tau_{n_{b}}\right]\frac{\phi^{\Psi(\tau_k) - \Psi(\tau_{k-1})}}{\Gamma\left[\Psi(\tau_k) - \Psi(\tau_{k-1})\right]}\lambda_k^{\Psi(\tau_k)-\Psi(\tau_{k-1}) - 1}
\end{align*}
where $\tau_0\equiv0$ and $\Gamma(x) = \int_0^\infty t^{x-1}e^{-t}dt$ is the gamma function. Even if $\Psi(\tau)$ is some nice known form, this density is nonstandard because the $\tau_k$'s enter the Gamma function. Some ideas for handlilng this step:
\begin{enumerate}
\item Draw each $\tau_k$ condtional on all the others using random walk Metropolis steps for $k=1,\dots,n_{b}$.
\item Approximate the density of $\tau_k$ sufficiently well and use independent Metropolis steps for $k=1,\dots,n_{b}$.
\item Approximate the density of $\tau_{1:n_{b}}$ sufficiently well and use an independent Metropolis step.
\item Use adaptive rejection sampling (ARS) to draw from the density of $\tau_k$.
\item Use ARMS --- ARS except with a Metropolis correction for densities which are not log concave (look it up).
\end{enumerate}

@@

\subsection{Regression parameters --- $\bm{\beta}$}
The full conditional posterior of $\bm{\beta}$ is closely connected to the functional form of the covariates in time. No matter the functional form, however, the density is complex and a Metropolis step is likely warranted. Recall that since we defined $\psi(t) = e^{\bm{z}(t)'\bm{\beta}}$ we have $\psi_k(\bm{\beta}) = \int_{\tau_{k-1}}^{\tau^k}e^{\bm{z}(t)'\bm{\beta}}dt$ --- in this subsection we write $\psi_k(\bm{\beta})$ to explicitly acknowledge that $\psi_k$ is a function of $\bm{\beta}$. Then the full conditional of $\bm{\beta}$ is
\begin{align*}
p(\bm{\beta}|\dots) \propto p(\bm{\beta})\prod_{k=1}^{n_b}\frac{\phi^{\psi_k(\bm{\beta})}\lambda_k^{\psi_k(\bm{\beta}) - 1}}{\Gamma[\psi_k(\bm{\beta})]}
\end{align*}
where in the prior $\bm{\beta} \sim p(\bm{\beta})$ independent of the other parameters. This density is complex and nonstandard since $\psi_k(\bm{\beta})$ enters the gamma function. A Metropolis step is likely warranted here as well.

@@

Otherwise, draw $\eta_k \sim Poi(\lambda_k)1(\eta_k \ge M_k - N_{k-1})$ and set $N_k = N_{k-1} + \eta_k$.
\end{enumerate}
\subsection{Block size caps --- $\bar{D}_k$}
Since the caps are independent discrete uniform in the prior, they are independent in their full conditional and for each block $k$, there are two possibilities:
\begin{enumerate}
\item When $N_k > M_k$, $D_k \le \bar{D}_k < D_k + d_{M_k+1}$.
\item When $N_k = M_k$, $D_k \le \bar{D}_k$.
\end{enumerate}
In the first case this should completely determine the value of $\bar{D}_k$ since transaction sizes are much smaller than the differences between possible caps. In both cases we draw $\bar{D}_k$ from a discrete uniform distribution over the set of possible caps given the relevant constraint above --- sometimes that distribution puts all of the mass on a single value.

\subsection{Transaction data parameters --- $\alpha_d$ and $\beta_d$}
The full conditional distribution of $(\alpha_d,\beta_d)$ is 
\begin{align*}
p(\alpha_d,\beta_d|...) \propto p(\alpha_d)p(\beta_d) \prod_{i=1}^{n_d}\frac{\beta_d^{\alpha_d}}{\Gamma(\alpha_d)}d_i^{\alpha_d - 1}e^{-\beta_dd_i}.
\end{align*}
When $\beta_d \sim G(a_\beta, b_\beta)$ in the prior, its full conditional becomes
\begin{align*}
p(\beta_d|...) \propto \beta_d^{\alpha_dn_d + a_{\beta} - 1}e^{-\beta_d\left(\sum_{i=1}^{n_d}d_i + b_{\beta}\right)},
\end{align*}
which is another gamma distribution. The full conditional of $\alpha_d$ is complex no matter what its prior is since $\alpha_d$ enters the gamma function, and has the form
\begin{align*}
p(\alpha_d|...) &\propto p(\alpha_d)\frac{\left(\beta_d^{n_d}\prod_{i=1}^{n_d}d_i\right)^{\alpha_d}}{\Gamma(\alpha_d)^{n_d}}\\
\intertext{which in the case of a gamma prior on $\alpha_d$ is}
p(\alpha_d|...) &\propto \alpha_d^{a_\alpha - 1}\frac{\left(e^{-b_\alpha}\beta_d^{n_d}\prod_{i=1}^{n_d}d_i\right)^{\alpha_d}}{\Gamma(\alpha_d)^{n_d}}.
\end{align*}
This density can be drawn from using rejection sampling or an independence Metropolis step, e.g. ARS or ARMS.
\bibliographystyle{apalike} 
\bibliography{../paper/btc}
